Antigua-Barbuda Fish Follows August 2017

### 2017-12-08 ABB_Analysis Tasks.Rmd

---

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.scar,int.con,int.hetscar,int.dam)
```

>>>>

---
Files: sumID2.csv, site.sum.csv

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.scar,int.con,int.hetscar,int.dam)
```

### 2017-12-08 ABB_Benthic.Rmd

### 1.1 Load packages
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# set up
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(plyr)
library(tidyr)
library(ggplot2)
library(grid)
library(stringr)
library(reshape2)
library(data.table)
```
### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/")
benthic<-read.csv("benthic.csv", stringsAsFactors=FALSE)
rugosity<-read.csv("rugosity.csv", stringsAsFactors=FALSE)
```

>>>>

### 1.1 Load packages
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# set up
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(plyr)
library(tidyr)
library(ggplot2)
library(grid)
library(stringr)
library(reshape2)
library(data.table)
```
### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/data")
benthic<-read.csv("benthic_anu.csv", stringsAsFactors=FALSE)
rugosity<-read.csv("rugosity.csv", stringsAsFactors=FALSE)
```

### 2017-12-08 ABB_Follows.Rmd

---
title: "Antigua-Barbuda Fish Follows August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---
```{r,message=FALSE,warning=FALSE}
# set up
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(tidyr)
library(ggplot2)
library(grid)
library(reshape2)
library(data.table)
```

```{r, warning=FALSE, message= FALSE}
# import data
setwd("~/github/Scarid-Behavior/data")
follows <- read.csv("follows.csv", stringsAsFactors=FALSE) 
follows <- follows %>% filter(Date != "" & Date !="3/26/17") %>% select(-Activity.Code) %>% unite("fishID",c(Site,Diver,Date,Time,Follow.Number),sep=".", remove=FALSE)

@@

# create df with single row for all fish followed
sumID <- follows %>% group_by(fishID, Date, Island, Site, Species, Species.Code, Phase, Length.cm, Weight.g, Follow.sec) %>% dplyr::summarize()
# summarize sample size by site, species, phase
n.sspph <- group_by(sumID, Site, Species, Phase) %>% dplyr::summarize(n=n())
```

# Grazing behavior

@@

## Grazing rates
```{r, warning=FALSE, message= FALSE}
# bites/second during total follow time and during grazing forays (including G1 instances)
bites <- follows %>% group_by(fishID) %>% dplyr::summarise(Bites = sum(Bites))
sumID <- left_join(sumID, bites, by="fishID") %>% mutate(BR=Bites/g.sec,FR=Bites/Follow.sec*60*60,TDB=Bites/Follow.sec*60*60*10.5)
sumID$BR[is.na(sumID$BR)] <- 0

@@

# Competitive interactions

## Summary of interactions per fish
```{r}
int.tot <- follows %>% filter(Initiator>=1) %>% group_by(fishID) %>% dplyr::summarise(int.tot=n())
int.scar <- follows %>% filter(Initiator>=1) %>% group_by(fishID) %>% filter(Competitor.Code=="stop" | Competitor.Code=="qup" | Competitor.Code=="rbp" | Competitor.Code=="prp" | Competitor.Code=="strp" | Competitor.Code=="rtp" | Competitor.Code=="ytp") %>% dplyr::summarise(int.scar=n())# total interaction instances
int.con <- follows %>% filter(Initiator>=1) %>% filter(Species==Competitor.Species) %>% group_by(fishID) %>% dplyr::summarise(int.con=n()) # interactions with conspecifics
int.hetscar <- follows %>% filter(Initiator>=1) %>% filter(Competitor.Code=="stop" | Competitor.Code=="qup" | Competitor.Code=="rbp") %>% filter(Species!=Competitor.Species) %>% group_by(fishID) %>% dplyr::summarise(int.hetscar=n()) # interactions with other scarid species
int.dam <- follows %>% filter(Initiator>=1) %>% filter(Competitor.Code=="dam") %>% group_by(fishID) %>% dplyr::summarise(int.dam=n()) # interactions with damselfish

@@

## Matrix of interactions by species
*eventually breakdown by site? region?*
```{r}
stop <- follows %>% filter(Species.Code=="stop") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(stop=n())
qup <- follows %>% filter(Species.Code=="qup") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(qup=n())
rbp <- follows %>% filter(Species.Code=="rbp") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(rbp=n())

@@

# Summarizing data by site, species, and phase
Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
```{r,warning=FALSE, message= FALSE}
# site averages
fol.site <- sumID %>% group_by(Site,Island) %>% dplyr::summarize(n=n(),g.frac.mean=mean(g.frac),g.frac.se=sd(g.frac)/sqrt(n),BR.mean=mean(BR,na.rm=TRUE),BR.se=sd(BR,na.rm=TRUE)/sqrt(n),FR.mean=mean(FR),FR.se=sd(FR)/sqrt(n),for.dur.mean=mean(for.dur,na.rm=TRUE),for.dur.se=sd(for.dur,na.rm=TRUE)/sqrt(n),for.bites.mean=mean(for.bites,na.rm=TRUE),for.bites.se=sd(for.bites,na.rm=TRUE)/sqrt(n),int.tot.mean=mean(int.tot),int.tot.se=sd(int.tot)/sqrt(n), int.scar.mean=mean(int.scar),int.scar.se=sd(int.scar)/sqrt(n), int.con.mean=mean(int.con),int.con.se=sd(int.con)/sqrt(n), int.hetscar.mean=mean(int.hetscar),int.hetscar.se=sd(int.hetscar)/sqrt(n), int.dam.mean=mean(int.dam),int.dam.se=sd(int.dam)/sqrt(n)) %>% select(-n)

@@

## Export csvs for integration with site data
```{r,message=FALSE,warning=FALSE}
# write.csv(sumID,file="sumID.csv")
# write.csv(fol.sp.site,file="fol.sp.site.csv")
# write.csv(fol.sp.ph.site,file="fol.sp.ph.site.csv")
```

### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.fish<-read.csv("site.fish.csv", stringsAsFactors=FALSE)%>% select(-X)
site.benthic<-read.csv("site.benthic.csv", stringsAsFactors=FALSE)%>% select(-X)
site.human<-read.csv("site.human.csv", stringsAsFactors=FALSE)
species.site<-read.csv("species.site.csv",stringsAsFactors = FALSE)%>% filter(Species=="Sparisoma viride" | Species=="Scarus vetula" | Species=="Sparisoma aurofrenatum") %>% select(-X,-Family,-n,-L.n)
names(species.site)[3:8] <- c("conspec.BM","conspec.BM.se","conspec.DEN","conspec.DEN.se","conspec.L","conspec.L.se")
```
### 1.3 Join datasets, print .csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# integrate all site-level variables
site.sum <- site.follows %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site") 
# then join site-level data to sumID dataframe
sumID2 <- sumID %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site")
# calculate conspecifc scarid variable to add to sumID2 (based on fish site/species)
sumID2 <- sumID2 %>% left_join(species.site,by=c("Site","Species"))

setwd("~/github/Scarid-Behavior/data")
write.csv(site.sum,file="site.sum.csv")
write.csv(sumID2,file="sumID2.csv")
```

>>>>

---
title: "Antigua-Barbuda Fish Follows August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---
# Antigua + Barbuda Fish Data
*required files: follows.csv*

## Setup

```{r,message=FALSE,warning=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(tidyr)
library(ggplot2)
library(grid)
library(reshape2)
library(data.table)
```

```{r, warning=FALSE, message= FALSE}
follows <- read.csv("follows.csv", stringsAsFactors=FALSE) 
follows <- follows %>% filter(Date != "" & Date !="3/26/17") %>% select(-Activity.Code) %>% unite("fishID",c(Site,Diver,Date,Time,Follow.Number),sep=".", remove=FALSE)

@@

# create df with single row for all fish followed
sumID <- follows %>% group_by(fishID, Date, Island, Site, Species, Species.Code, Phase, Length.cm, Weight.g, Follow.sec) %>% dplyr::summarize()
# summarize sample size by site, species, phase
n.sspph <- group_by(sumID, Site, Species, Phase) %>% dplyr::summarize(n=n())

```

@@

## Bite rates (bites/sec) overall and while grazing
```{r, warning=FALSE, message= FALSE}
# bites/second during total follow time and during grazing forays (including G1 instances)
bites <- follows %>% group_by(fishID) %>% dplyr::summarise(Bites = sum(Bites))
sumID <- left_join(sumID, bites, by="fishID") %>% mutate(BR=Bites/g.sec,FR=Bites/Follow.sec*60*60,TDB=Bites/Follow.sec*60*60*10.5)
sumID$BR[is.na(sumID$BR)] <- 0

@@

## Competitive interactions

### Summary of interactions per fish
```{r}
int.tot <- follows %>% filter(Initiator>=1) %>% group_by(fishID) %>% dplyr::summarise(int.tot=n())
int.scar <- follows %>% filter(Initiator>=1) %>% group_by(fishID) %>% filter(Competitor.Code=="stop" | Competitor.Code=="qup" | Competitor.Code=="rbp" | Competitor.Code=="prp" | Competitor.Code=="strp" | Competitor.Code=="rtp" | Competitor.Code=="ytp") %>% dplyr::summarise(int.scar=n())# total interaction instances
int.con <- follows %>% filter(Initiator>=1) %>% filter(Species==Competitor.Species) %>% group_by(fishID) %>% dplyr::summarise(int.con=n()) # interactions with conspecifics
int.hetscar <- follows %>% filter(Initiator>=1) %>% filter(Competitor.Code=="stop" | Competitor.Code=="qup" | Competitor.Code=="rbp") %>% filter(Species!=Competitor.Species) %>% group_by(fishID) %>% dplyr::summarise(int.hetscar=n()) # interactions with other scarid species
int.dam <- follows %>% filter(Initiator>=1) %>% filter(Competitor.Code=="dam") %>% group_by(fishID) %>% dplyr::summarise(int.dam=n()) # interactions with damselfish

@@

### Matrix of interactions by species
*eventually breakdown by site? region?*
```{r}
stop <- follows %>% filter(Species.Code=="stop") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(stop=n())
qup <- follows %>% filter(Species.Code=="qup") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(qup=n())
rbp <- follows %>% filter(Species.Code=="rbp") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(rbp=n())

@@

## Summarizing data by site, species, and phase
Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
```{r,warning=FALSE, message= FALSE}
# site averages
fol.site <- sumID %>% group_by(Site,Island) %>% dplyr::summarize(n=n(),g.frac.mean=mean(g.frac),g.frac.se=sd(g.frac)/sqrt(n),BR.mean=mean(BR,na.rm=TRUE),BR.se=sd(BR,na.rm=TRUE)/sqrt(n),FR.mean=mean(FR),FR.se=sd(FR)/sqrt(n),for.dur.mean=mean(for.dur,na.rm=TRUE),for.dur.se=sd(for.dur,na.rm=TRUE)/sqrt(n),for.bites.mean=mean(for.bites,na.rm=TRUE),for.bites.se=sd(for.bites,na.rm=TRUE)/sqrt(n),int.tot.mean=mean(int.tot),int.tot.se=sd(int.tot)/sqrt(n), int.scar.mean=mean(int.scar),int.scar.se=sd(int.scar)/sqrt(n), int.con.mean=mean(int.con),int.con.se=sd(int.con)/sqrt(n), int.hetscar.mean=mean(int.hetscar),int.hetscar.se=sd(int.hetscar)/sqrt(n), int.dam.mean=mean(int.dam),int.dam.se=sd(int.dam)/sqrt(n)) %>% select(-n)

@@

## Export csvs for integration with follow data
```{r,message=FALSE,warning=FALSE}
write.csv(site.follows,file="site.follows.csv")
write.csv(sumID,file="sumID.csv")
write.csv(fol.sp.site,file="fol.sp.site.csv")
write.csv(fol.sp.ph.site,file="fol.sp.ph.site.csv")
```
